---
title: "manuscript"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "15/02/2021"
output: html_document
---
```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(lmerTest)
library(lme4)
library(helprs)
library(papaja)
source("../functions/cdz-preprocess.R")
source("functions/reporting.R")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

CDZ_COLORS <- c("#63873a", "#f3f7e9", "#25aae1", "#81c6e2")

df_results <- read.table("../run2-results-supermarket-sessions.csv", sep = ";",
                         header = TRUE)

df_demographics <- read.table("../demographics.csv", sep = ";", header = TRUE)
df_ids <- read.table("../information.csv", header = TRUE, sep = ";")

df_sessions <- read.table("../sessions.csv", header = TRUE, sep = ";")

## RBANS ----
df_rbans <- read.table("../rbans.csv", sep = ";", header = TRUE) %>%
  add_session_to_rbans(df_demographics) %>%
  add_summaries_to_rbans()

good_participants <- df_rbans %>%
  group_by(rbansid) %>%
  filter(!is.na(testpameti)) %>%
  summarise(n = n()) %>%
  filter(n >= 2) %>%
  pull(rbansid)

df_rbans <- df_rbans %>%
  filter(rbansid %in% good_participants)

df_rbans_long <- df_rbans %>%
    pivot_longer(-c(rbansid, trenink, diagnosis, first_training, session),
                 names_to = "subtest", values_to = "score") %>%
  group_by(session, subtest) %>%
  mutate(z_score = scale(score)[,1])

df_rbans_long <- df_rbans_long %>%
  group_by(trenink, first_training, subtest) %>%
  summarise(group_avg = mean(score, na.rm = TRUE),
            group_z_avg = mean(z_score, na.rm = TRUE),
            .groups="drop") %>%
  right_join(df_rbans_long, by = c("trenink", "first_training", "subtest"))

## SUPERMARKET ------
df_results <- df_demographics %>%
  left_join(df_ids, by = "rbansid") %>%
  select(rbansid, vrid) %>% 
  filter(!is.na(rbansid) & !is.na(vrid)) %>%
  right_join(df_results, by = c("vrid" = "participant")) %>%
  mutate(participant = vrid) %>%
  select(-vrid)

df_results <- df_results %>%
  mutate(total_picked_items = n_correct_items + n_extra_items,
         correct_ratio = n_correct_items/n_items,
         extra_ratio = n_extra_items/n_items) %>%
  group_by(participant, session) %>%
  mutate(n_trials = n(), 
         max_difficulty = max(n_items)) %>%
  ungroup()

# First session was supposedly weird, and only few people did more than 12
df_results <- filter(df_results, session <= 12, session > 1)
df_results <- filter(df_results, rbansid %in% good_participants)
df_results_remediation <- filter(df_results, type == "remediation")
df_results_test <- filter(df_results, type == "test")

## Questionnaire ----
df_questionnaire <- read.table("../questionnaire.csv", sep = ";", header = TRUE) %>%
  mutate_at(vars(starts_with(c("vr", "paper", "supermarket"))), 
            function(x){ifelse(x == 6, NA, x)}) %>%
  left_join(df_demographics, c("rbansid" = "rbansid"))

df_demographics_good <- df_demographics %>%
  filter(rbansid %in% good_participants)
```

## Methods
Due to missing standardization procedures for calculating RBANS IQ scores in version C and an arguable difference between the three versions, we have opted to use the raw scores and not the standard metrics in RBANS related calculations. We normalized our raw scores in each RBANS domain for each version separately. As our models take RBANS version into account to control for potential raw score differences between different versions and because of the normalization procedure, we consider this approach in our scenario to be more statistically powerful.

## Demographics

A total of `r nrow(df_demographics)` participants were recruited. Only those participants who completed at least one intervention session were included in the subsequent analyses: `r nrow(df_demographics_good)` in total, `r table(df_demographics_good$gender)['male']` male (`r mean_and_sd_report(df_demographics[df_demographics$gender == "male", "age"])`), female `r table(df_demographics_good$gender)['female']` (`r mean_and_sd_report(df_demographics[df_demographics$gender == "female", "age"])`).

Demografická tabulka s průměry délky onemocnění, věkem, vzděláním atp - případně vytvořím sama, můžeš třeba jen vysypat ta data (a nebo to taky udělám)

```{r}
df_demographics_good %>%
  select(age, ilness_duration_years) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  knitr::kable(digits = 3)

df_demographics_good %>%
  count(diagnosis) %>%
  knitr::kable()

df_demographics_good %>%
  count(education) %>%
  knitr::kable()
```

### Different diagnoses

We have observed that patients suffering from schizophrenia had significantly worse baseline RBANS performance than patients with MDD (see table XXX). This was one of the reasons we opted for the linear mixed modeling in our subsequent analyses as they should be able to control for the effect of diagnosis.

```{r}
df_demographics_good %>%
  mutate(diagnosis = sapply(diagnosis, function(x){
    if(grepl("F3.*", x)) return("affective")
    if(grepl("F2.*", x)) return("schizophrenia")
  })) %>%
  count(diagnosis, first_training) %>%
  pivot_wider(names_from = diagnosis, values_from = n) %>%
  knitr::kable()

df_rbans_long %>%
  filter(grepl("dim_", subtest)) %>%
  ggplot(aes(subtest, z_score, fill=diagnosis)) +
  geom_boxplot()

df_diag_difference <- df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  group_by(subtest) %>%
  summarise(t_test = apa_print(t.test(score ~ diagnosis))$statistic, 
            .groups = "drop")

df_rbans_long %>%
  filter(grepl("iq_", subtest), session == 1) %>%
  group_by(subtest, diagnosis) %>%
  summarise(m = mean_and_sd_report(score, type = "parentheses")) %>%
  pivot_wider(names_from = diagnosis, values_from = m) %>%
  right_join(df_diag_difference, by = "subtest") %>%
  knitr::kable(digits = 3)

df_rbans %>%
  group_by()

```

Adherence rate – potřebujeme info o tom, kolik lidí bylo zahrnuto do studie (splnili kritéria), kolik lidí odpadlo v průběhu VR  nebo TP, kolik lidí odmítlo začít VR/TP - já bych doplnila důvody odpadu

## Questionnaire

```{r}
df_questionnaire_long <- df_questionnaire %>%
  pivot_longer(cols = c(starts_with(c("vr", "paper")), -ends_with("feedback")),
               names_to = "questionnaire_section", values_to = "score") %>%
  separate(questionnaire_section, sep = "_",
           into = c("question_focus", "question_area"),
           extra="merge")

df_temp <- df_questionnaire_long %>%
  group_by(question_area, question_focus) %>%
  summarise(mean = mean_and_sd_report(score, type = "parentheses")) %>%
  pivot_wider(names_from = question_focus, values_from = mean)

df_questionnaire_long %>%
  group_by(question_area) %>%
  summarise(t.test = apa_print(t.test(score~question_focus))$statistic) %>%
  right_join(df_temp, by = "question_area") %>%
  select(question_area, paper, vr, t.test) %>%
  knitr::kable(digits = 3, caption = "")

n_vr <- sum(!is.na(df_questionnaire$vr_fun))
n_paper <- sum(!is.na(df_questionnaire$paper_fun))

t_questionnaire_repeat <- t.test(df_questionnaire$vr_would_repeat,
                                 df_questionnaire$paper_would_repeat)
```

The questionnaire aimed at assessing patients’ subjective perception of the intervention in terms of difficulty, fun or improvement indicate no significant difference between the standard treatment and the VEs program in either of the five categories (see table XXX). `r n_vr` participants answered the questions about the VR portion of the program and `r n_paper` about the standard treatment. Although the patients seem to give a slight edge to the paper tests in terms of repeatability `r apa_print(t_questionnaire_repeat)$statistic`.

## Supermarket

```{r}
df_results_remediation %>%
  group_by(session, n_items) %>%
  summarise(correct_ratio = mean_and_sd_report(correct_ratio, type="parentheses"),
            time = mean_and_sd_report(results_time, type="parentheses"),
            trajectory = mean_and_sd_report(results_trajectory, type="parentheses")) %>%
  filter(session == 2 | session == 12) %>%
  knitr::kable(caption = "Select what might be relevant or what to aggregate together")
```

```{r}
df_results_remediation %>%
  select(rbansid, session, max_difficulty) %>%
  distinct() %>%
  ggplot(aes(session, max_difficulty)) + 
    geom_jitter(height = 0, width = 0.15) +
    geom_smooth(method = "lm", fill = CDZ_COLORS[4], color = CDZ_COLORS[3]) +
    scale_x_continuous(breaks = seq(2,12,1)) + 
    theme_apa() +
    labs(y = "Maximum difficulty achieved", x = "Session")
ggsave("figures/fig1-session-difficulty.png",  width = 8, height = 6)
```

```{r mixed models vSST results}
lmer_diff_session <- df_results_remediation %>%
  select(rbansid, session, max_difficulty) %>%
  distinct() %>%
  lmer(max_difficulty ~ session + (1|rbansid), data = .)

lmer_time_session_trial <- lmer(results_time ~ session + n_items + (1|participant), 
                                data = df_results_remediation)
lmer_trajectory_session_trial <- lmer(results_trajectory ~ session + n_items + (1|participant), 
                                      data = df_results_remediation)
lmer_correct_session_6_9 <- lmer(correct_ratio ~ session + n_items + (1|participant), 
                                 data = filter(df_results_remediation, n_items >= 6))
```

Using mixed effect modeling we looked at how patients learned the task as the cognitive training progressed. We observed that participants were continuously able to improve and proceed to more difficult trials as the training progressed `r report_mixed_effect(lmer_diff_session, "session")`. 

As the trajectory and time of trials is dependent both on learning as well as task difficulty (more items take longer to pick up), we modeled the trial time and trajectory with both session and trial difficulty as a predictor. We  observed that participants improved in both these measures as cognitive training progressed, seeing decrease in trial trajectories (`r report_mixed_effect(lmer_trajectory_session_trial, "session")`) and trial times (`r report_mixed_effect(lmer_time_session_trial, "session")`) with each new session. 

One of the other metric we explored previously and found to be indicative of cognitive performance (PLECHATA CS PSY) was the vSST correct ratio. As the correct ratio is dependent on task difficulty (more difficult trials lead to higher performance decline), and as we learned in our previous work that patients only start to struggle after 5 item difficulty, we only analyzed this metric in trials beginning with difficulty of six items. Modeling the correct ratio as a function of session and task difficulty, we observed that indeed the task difficulty significantly decreases the performance in this metric (`r report_mixed_effect(lmer_correct_session_6_9, "n_items")`), but number of completed sessions had no effect (`r report_mixed_effect(lmer_correct_session_6_9, "session")`). 

XXX - discussion?? Drop ? Patients did eventually improve in the vSST, as suggested by the upward trend in achieved difficulty with progressing sessions, and due to the nature of the task (higher difficulty is only triggered if the correct ratio of the current one is perfect) this must mean that patients must have had slightly improved in this metric as well, although this change was much more gradual and is "noised XXX" by the number of trial repetitions patients did.

## RBANS

```{r}
df_rbans %>%
  group_by(session, trenink) %>%
  summarise_at(vars(starts_with("iq_")), list(
            desc = ~mean_and_sd_report(.x, type = "parentheses"),
            n = ~sum(!is.na(.x)))) %>%
  mutate(number = iq_kratkodobapamet_n) %>%
  select(-ends_with("_n")) %>%
  rename_with(~gsub("iq_(.*)_desc", "\\1", .x)) %>%
  knitr::kable()
```


```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(fac_subtest, score, fill = trenink)) +
    geom_boxplot() +
    facet_wrap(~session,  strip.position = "bottom") +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
      "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
      "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program"),
                        values = CDZ_COLORS) +
    theme_apa() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
    labs(x = "", y = "RBANS index score",
         title = "Index scores of RBANS cognitive domains",
         fill = "")

ggsave("figures/fig2-rbans-iq.png",  width = 8, height = 6)

df_rbans_long %>%
  filter(subtest == "dim_rec") %>%
  ggplot(aes(score, fill = factor(session))) + 
  geom_histogram() + 
  facet_wrap(~trenink)
```

```{r}
df_rbans_no_baseline <- filter(df_rbans, session != 1)
```

```{r}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname, session > 1) %>%
    lmer(z_score ~ trenink + factor(session) + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- temp_mixed_func("dim_rec")
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 2)})
knitr::kable(as.data.frame(res),
             caption = "Trenink as a predictor of RBANS performance. 
             Coefficients are for VR in contrast to TP (negative = VR does worse)")
```


Using linear mixed effect modeling we explored the effect of the intervention as well as the session on the RBANS scores with the participant as a random effect to remove within subject differences in RBANS performance. We have not found any effect, neither positive nor negative, of intervention type (see table XXX). We have only found a negative effect of session on the normalized total RBANS score `r report_mixed_effect(lmer_celek, 3)` with people at the end of the study (third RBANS, both interventions completed) scoring marginally worse than in previous tests. This can be due to XXXXX

```{r rbans czech norms, eval=FALSE}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  ggplot(aes(subtest, score, fill = factor(session))) +
    geom_boxplot() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position  = "top", legend.direction = "horizontal")

df_rbans_long %>%
  filter(grepl("dim_", subtest)) %>%
  ggplot(aes(subtest, score, fill = factor(session))) +
    geom_boxplot() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position  = "top", legend.direction = "horizontal")

df_rbans_long %>%
  filter(!grepl("(iq|dim)_", subtest)) %>%
  ggplot(aes(subtest, score, fill = factor(session))) +
    geom_boxplot() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position  = "top", legend.direction = "horizontal")

temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname, session > 1) %>%
    lmer(score ~ trenink + factor(session) + (1|rbansid), data = .)
  return(out)
}

temp_glm_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname) %>%
    glm(score ~ session, data = .)
  return(out)
}

summary(temp_mixed_func("iq_celek"))
summary(temp_glm_func("iq_rec"))
```

```{r session model explanation, eval = FALSE}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname) %>%
    lmer(z_score ~ factor(session) + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <-temp_mixed_func("dim_pozornost")
lmer_rec <- temp_mixed_func("dim_rec")
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 2)})
knitr::kable(as.data.frame(res),
             caption = "Trenink as a predictor of RBANS performance.
             Coefficients are for VR in contrast to TP (negative = VR does worse)")

df_rbans_long %>%
  filter(subtest == "dim_celek") %>%
  mutate(session = paste0("session ", session - 1)) %>%
  ggplot(aes(session, z_score, fill = trenink)) +
    geom_boxplot() +
    geom_jitter(position = position_dodge(width = 0.5)) +
    scale_fill_discrete(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program")) +
    geom_hline(yintercept = 0.07) +
    theme_apa() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
    labs(x = "", y = "Standardized score",
         title = "Standardized scores of RBANS cognitive domains",
         fill = "")

df_rbans_long %>%
  filter(subtest == "dim_celek") %>%
  ggplot(aes(session, z_score, group = rbansid)) + 
    geom_line() + geom_point(aes(color = trenink), size = 3)

df_rbans_long %>%
  filter(subtest == "dim_celek") %>%
  group_by(session) %>%
  summarise(mean_and_sd_report(score))
```

## Supplementary

```{r diagnosis model, eval = TRUE}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname, session > 1) %>%
    lmer(z_score ~ trenink + diagnosis + factor(session) + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- temp_mixed_func("dim_rec")
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 3)})
knitr::kable(as.data.frame(res),
             caption = "Diagnosis as a predictor of RBANS performance. 
             Coefficients are for schizophrenia in contrast to MDD (negative = schizophrenia does worse)")
```
```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(subtest, score, fill = trenink)) +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
    "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
    "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    geom_boxplot() + 
    theme_apa() +
    scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program"),
                        values = CDZ_COLORS) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
      labs(x = "", y = "RBANS index score",
         title = "Index scores of RBANS cognitive domains after each intervention",
         fill = "")

ggsave("figures/fig3-rbans-iq.png",  width = 8, height = 6)
```

## Tesuceni

Looking at the RBANS improvement as a function of a learning baseline bring significant results UNTIL we consider that the RBANS dimensions are not independent. In other works, person doing well on one often does well in others. We get significant effects of baseline tests in other tests as well (e.g. good baseline performance in orientace primky predicts good performance in ) When we control for the rbans dimension scores by their overall score in the session, the effect of the learning disappears. 

```{r effect of testuceni, eval=FALSE}
temp_mixed_func <- function(varname){
  out <- df_temp %>%
    filter(subtest == varname, session > 1) %>%
    lmer(score ~ total_score + learning_baseline + factor(session) + (1|rbansid), data = .)
  return(out)
}

df_temp2 <- baseline_rbans(df_rbans) %>%
  left_join(select(df_rbans, rbansid, learning = testuceni), by = "rbansid")

summary(lmer(dim_celek ~ learning + factor(session) + (1|rbansid), data = df_temp2))

df_temp <- df_rbans_long %>%
  arrange(rbansid, session) %>%
  group_by(rbansid) %>%
  mutate(learning_baseline = score[subtest == "testuceni"][1]) %>%
  group_by(session, .add = TRUE) %>%
  mutate(total_score = score[subtest == "dim_celek"][1]) %>%
  ungroup()

summary(temp_mixed_func("dim_kratkodobapamet"))
summary(temp_mixed_func("dim_dlouhodobapamet"))
summary(temp_mixed_func("dim_pozornost"))
summary(temp_mixed_func("dim_rec"))
summary(temp_mixed_func("dim_vizuoprostor"))

dims <- c("dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")

res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 3)})
knitr::kable(as.data.frame(res),
             caption = "Testuceni in baseline as a predictor of rbans improvement controlled by rbans version (session) and total score")
```
