---
title: "Supermarket ouptut"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lmerTest)
library(lme4)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE)
df_results <- read.table("../run2-results-supermarket-sessions.csv", sep = ";",
                         header = TRUE)

df_demographics <- read.table("../demographics.csv", sep = ";", header = TRUE)
df_ids <- read.table("../information.csv", header = TRUE, sep = ";")

# Combine with demographics
df_join <- df_demographics %>%
  left_join(df_ids, by = "rbansid") %>%
  select(rbansid, vrid) %>% 
  filter(!is.na(rbansid) & !is.na(vrid))

df_results <- df_results %>%
  left_join(df_join, by = c("participant" = "vrid"))
```

```{r}
df_results <- df_results %>%
  mutate(total_picked_items = n_correct_items + n_extra_items,
         correct_ratio = n_correct_items/n_items,
         extra_ratio = n_extra_items/n_items) %>%
  group_by(participant, session) %>%
  mutate(n_trials = n()) %>%
  ungroup()

# First session was supposedly weird, and only few people did more than 12
df_results <- filter(df_results, session <= 12, session > 1)

df_results_remediation <- filter(df_results, type == "remediation")
df_results_test <- filter(df_results, type == "test")
```


### Effect of session and number of items on trial time

```{r}
df_results_remediation %>%
  filter(results_time < 400) %>%
  ggplot(aes(session, results_time)) +
    geom_jitter(height = 0.1, width = 0.2) +
    scale_x_continuous(breaks = 2:12) +
    geom_smooth(method = "lm") + 
    facet_wrap(~n_items)
```

```{r}
df_results_remediation %>%
  group_by(session, n_items) %>%
  summarise(results_time = mean(results_time)) %>%
  ggplot(aes(session, y = results_time)) +
    geom_line() +
    facet_wrap(~n_items) +
    scale_x_continuous(breaks = 2:12) +
    labs(y = "average trial time")
```


```{r}
lmer(results_time ~ session + n_items + (session | participant),
  df_results_remediation) %>%
  summary()
```
### Effect of session and number of items on correct ratio

```{r}
df_results_remediation %>%
  group_by(participant, session) %>%
  summarise(correct_ratio = mean(correct_ratio)) %>%
  ungroup() %>%
  ggplot(aes(session, correct_ratio)) +
    geom_line(aes(color=participant))
```


```{r}
df_results_remediation %>%
  group_by(session, participant) %>%
  summarise(correct_ratio = mean(correct_ratio), .groups = "drop_last") %>%
  summarise(correct_ratio = mean(correct_ratio)) %>%
  ggplot(aes(session, correct_ratio)) +
    geom_line() + scale_y_continuous(limits = c(0, 1))
```


```{r}
df_results_remediation %>%
  ggplot(aes(session, correct_ratio)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_wrap(~n_items)

df_results_remediation %>%
  filter(n_items > 6) %>%
  ggplot(aes(session, correct_ratio, color = factor(n_items))) +
  geom_jitter(width = 0.25) +
  geom_smooth(aes(color = "all"), method = "lm")
```

```{r}
lmer(correct_ratio ~ session + n_items + (session | participant),
  df_results_remediation) %>%
  summary()
```

AS we could expect, the ratio stays the same. But participants generally achieve better times and higher difficulties, so that is not necessarily a problem.

### Effect of session and number of items on trial trajectory

```{r}
df_results_remediation %>%
  ggplot(aes(session, results_trajectory)) +
    geom_jitter(width = 0.1, height = 0.3) +
    scale_y_continuous(limits = c(0,750)) +
    geom_smooth(method = "lm") + 
    facet_wrap(~n_items) +
  labs(y = "trial trajectory")
```

```{r}
lmer(results_trajectory ~ session + n_items + (session | participant),
  df_results_remediation) %>%
  summary()
```


### Effect of session on maximum achieved difficulty
```{r}
df_results_remediation %>%
  group_by(session, participant) %>%
  summarise(max_difficulty = max(n_items), .groups="drop") %>%
  ggplot(aes(session, y = max_difficulty, color = participant)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~participant) +
    guides(color = FALSE) +
    scale_x_continuous(breaks = 2:12) +
    labs(y = "Maximum achieved difficulty")
```

```{r}
df_results_remediation %>%
  group_by(session, participant) %>%
  summarise(max_difficulty = max(n_items)) %>%
  ggplot(aes(session, max_difficulty)) +
    geom_jitter(aes(color=participant), height = 0.1, width = 0.2) +
    geom_smooth(method = "lm") +
    scale_x_continuous(breaks = 2:12) +
    scale_y_continuous(breaks = 1:15) +
    labs(y="maximum achieved difficulty") + guides(color=FALSE)
```

```{r}
df_results_remediation %>%
  group_by(session, participant) %>%
  summarize(max_difficulty = max(n_items), .groups = "drop") %>%
  lmer(max_difficulty ~ session  + (session|participant), data = .) %>%
    VarCorr()
```

```{r}
df_results_remediation %>%
  group_by(session, participant) %>%
  summarize(max_difficulty = max(n_items), .groups = "drop") %>%
  lmer(max_difficulty ~ session  + (session | participant), data = .) %>%
    summary()
```
